---
title: 'Bios 6301: Homework 2'
output: pdf_document
author: Zi Ye
---

**Grade: 52/50**

*(informally) Due Tuesday, 20 September, 1:00 PM*

50 points total.

This assignment won't be submitted until we've covered Rmarkdown.
Create R chunks for each question and insert your R code appropriately.
Check your output by using the `Knit PDF` button in RStudio.

1. **Working with data** In the `datasets` folder on the course GitHub repo, you will find a file called `cancer.csv`, which is a dataset in comma-separated values (csv) format. This is a large cancer incidence dataset that summarizes the incidence of different cancers for various subgroups. (18 points)

    1. Load the data set into R and make it a data frame called `cancer.df`. (2 points)
  
    ```{r}
    #setwd('~/Downloads/Bios6301-master/datasets')
    x <- data.frame(read.csv('cancer.csv'))
    ```
    2. Determine the number of rows and columns in the data frame. (2)
    ```{r}
    nrow(x)
    ncol(x)
    ```
    3. Extract the names of the columns in `cancer.df`. (2)
    ```{r}
    names(x)
    ```
    4. Report the value of the 3000th row in column 6. (2)
    ```{r}
    x[3000,6]
    ```
    5. Report the contents of the 172nd row. (2)
    ```{r}
    x[172,]
    ```
    6. Create a new column that is the incidence *rate* (per 100,000) for each row.(3)
    ```{r}
    incirate <- x$incidence/100000
    x <- data.frame(x, incirate)
    ```
    
**JC Grading - 1**
For incidence rate above should be incidence / population * 100000

    
    7. How many subgroups (rows) have a zero incidence rate? (2)
    ```{r}
    nrow(x[incirate==0,])
    ```
    8. Find the subgroup with the highest incidence rate.(3)
    ```{r}
    x[incirate==max(x$incirate),]
    ```
    
**JC Grading - 1**
syntax is fine but answer is incorrect b/c of how incidence rate was calculated
    
    
2. **Data types** (10 points)

    1. Create the following vector: `x <- c("5","12","7")`. Which of the following commands will produce an error message? For each command, Either explain why they should be errors, or explain the non-erroneous result. (4 points)
    
            max(x)
            sort(x)
            sum(x)
    
    **Since the number within quote mark will become characters, `x` is not a combination of numbers but characters. `sum(x)` returns error message since it is impossible to `sum()` characters. `max(x)` can be used on characters but it only takes the first letter in each element into account ('12' becomes '1'), and so as `sort(x)`.**
      
    2. For the next two commands, either explain their results, or why they should produce errors. (3 points)

            y <- c("5",7,12)
            y[2] + y[3]

    **It returns errors. The first element in `y` is a character. The `c()` function will switch all the elements to character if there are both numbers and characters. Therefore, `y[2]` becomes the character '7' and `y[3]` becomes the character '12', and it is impossible to add them together.**

    3. For the next two commands, either explain their results, or why they should produce errors. (3 points)

            z <- data.frame(z1="5",z2=7,z3=12)
            z[1,2] + z[1,3]
            
    **It returns 19. As opposed to `c()` function, `data.frame()` function switchs characters to numbers when both characters and numbers are included in the data frame. So the second and third elements can be added up to 19.**
    
3. **Data structures** Give R expressions that return the following matrices and vectors (*i.e.* do not construct them manually). (3 points each, 12 total)

    1. $(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)$
    ```{r}
    c(seq(8), seq(7, by=-1))
    ```
    
    2. $(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)$
    ```{r}
    c(1,rep(2,2),rep(3,3),rep(4,4),rep(5,5))
    ```

    3. $\begin{pmatrix}
      0 & 1 & 1  \\
      1 & 0 & 1  \\
      1 & 1 & 0  \\
    \end{pmatrix}$
    ```{r}
    m1 <- matrix(1, nrow=3, ncol=3)
    diag(m1) <- 0
    m1
    ```
    4. $\begin{pmatrix}
      1 & 2 & 3 & 4 \\
      1 & 4 & 9 & 16 \\
      1 & 8 & 27 & 64  \\
      1 & 16 & 81 & 256 \\
      1 & 32 & 243 & 1024  \\
    \end{pmatrix}$
    ```{r}
    m2 <- matrix(ncol=4, nrow=5, byrow=T)
    for (i in 1:4) {
    m2[,i]=i^(1:5)
    }
    m2
    ```
4. **Basic programming** (10 points)

    1. Let $h(x,n)=1+x+x^2+\ldots+x^n = \sum_{i=0}^n x^i$. Write an R program to calculate $h(x,n)$ using a `for` loop. (5 points)
    
                    h = 0
                    for (i in 0:n) {
                    h = h + x^i
                    }
                    h
     
    2. If we list all the natural numbers below 10 that are multiples of 3 or 5, we get 3, 5, 6 and 9. The sum of these multiples is 23. Write an R program to perform the following calculations. (5 points)

        1. Find the sum of all the multiples of 3 or 5 below 1,000. (3, [euler1])
        
    ```{r}
    x = 0
    for (i in 1:999) {
    if (i%%3==0 | i%%5==0) {
    x = x+i
    }
    }
    x
    ```
        2. Find the sum of all the multiples of 4 or 7 below 1,000,000. (2)
    ```{r}
    y = 0
    for (i in 1:(1e6-1)) {
    if (i%%4==0 | i%%7==0) {
    y = y+i
    }
    }
    y
    ```
    3. Each new term in the Fibonacci sequence is generated by adding the previous two terms. By starting with 1 and 2, the first 10 terms will be $(1, 2, 3, 5, 8, 13, 21, 34, 55, 89)$. Write an R program to calculate the sum of the first 15 even-valued terms. (5 bonus points, [euler2])
```{r warning=FALSE}
    z = c(1, 2)
    even = c(2)
    for (i in 3:1000) {
      z[i] = z[i-1] + z[i-2]
      if (z[i]%%2==0 & length(even) < 16) {
        even = c(even, z[i])
      }
    }
    sum(even)
```

**JC Bonus +4**
Included one extra term (the first term 2 was included before loop)


Some problems taken or inspired by projecteuler.

[euler1]:https://projecteuler.net/problem=1
[euler2]: https://projecteuler.net/problem=2
